Design and implementaiton of a live coding environment

### 2017-10-03 live/paper.tex
Properties of previews

@@

\section{Properties}
\newpage

>>>>

\newcounter{thc}
\newcounter{dfc}

\theoremstyle{plain}
\newtheorem{theorem}[thc]{Theorem}
\newtheorem{lemma}[thc]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[dfc]{Definition}

@@

\newpage
\section{Properties}

\subsection{Host language}

We require that the host language allows eliminating let bindings without changing the meaning
of the program. That is

\begin{definition}[Host language reduction]
The reduction $\rightsquigarrow$ reduces host language expression such that:
\begin{itemize}
\item  
Given an expression $e$ and $e'$ such that $e'$ is obtained from $e$
by replacing any sub-expression $\kvd{let}~x=e_1~\kvd{in}~e_2$ with the expression 
$e_2[x\leftarrow e_1]$, written as $e\rightsquigarrow_\ident{let}e'$, then if $e \rightsquigarrow e''$ then also $e' \rightsquigarrow e''$.

\item 
Compositionality, i.e. if $e \rightsquigarrow e'$ and $C[e] \rightsquigarrow e''$ then also
  $C[e'] \rightsquigarrow e''$.
\end{itemize}
\end{definition}

We don't care about non-termination here.
(so, does $\rightsquigarrow$ do small step or big step reduction?)

\subsection{Correctness of previews}
Eliminating let bindings does not change the result in the host language. It also yields the same
dependency graph.

\begin{lemma}[Let elimintion]
Given an expression $e_1$ such that $e_1 \rightsquigarrow_\ident{let} e_2$ and a lookup table $\Delta$
then if $\bndclr{v_1}, (V_1, E_1) = \ident{bind}_{\emptyset, \Delta}(e_1)$ and  
$\bndclr{v_2}, (V_2, E_2) = \ident{bind}_{\emptyset, \Delta}(e_2)$ then it holds that
$\bndclr{v_1} = \bndclr{v_2}$ and also $(V_1, E_1) = (V_2, E_2)$.
\end{lemma}

This simplifies the correctness proof as we only need to consider expressions not containing
let bindings (if there were let bindings, both the result and the preview would be equivalent
to one that would be obtained if the let bindings were elimninated)

Now, if we have any let-free expression, we show that the preview obtained via the dependency graph
is a value that is the same as the one we'd get by directly evaluating the expression.

\begin{lemma}[Let-free correctness]
Given an expression $e$ that has no free variables and does not contain let bindnigs, together with
a lookup table $\Delta$ obtained from any sequence of expressions according to Figure~\ref{fig:loop}, 
let $\bndclr{v}, (V, E) = \ident{bind}_{\emptyset, \Delta}(e)$. If $\bndclr{v}\Downarrow d$ 
over a graph $(V, E)$ then $d = p$ for some $p$ and $e \rightsquigarrow p$.
\end{lemma}
\begin{proof}
Every dependency graph node $\bndclr{v}$ corresponds to an expression $e$ with free variables
$FV(e)$. We obtain preview $\bndclr{v}\Downarrow d$ such that if $FV(e) = \emptyset$ then
$d = p$ for some $p$, otherwise, $d = \llbracket e' \rrbracket_{FV(e)}$.  
We show that either $e \rightsquigarrow p$ or that for any $C$ such that $FV(C[e]) = \emptyset$
it holds that $C[e] \rightsquigarrow e''$ and $C[e'] \rightsquigarrow e''$.
By induction over the structure of the evaluation:

\vspace{0.5em}\noindent\emph{(val)} $\bnd{val}(n) \Downarrow n$ and so $p=n$. The only
  \ident{bind} case that generates $\bnd{val}(n)$ in the absence of let bindings is~(4) and
  so $e = n$.

\vspace{0.5em}\noindent\emph{(mem-val)} $\bnd{mem}(m, s) \Downarrow p$. By induction, there
  are nodes $\bndclr{v_i}$ representing expressions $e_i$ and previews $\bndclr{v_i}\Downarrow p_i$
  such that $e_i \rightsquigarrow p_i$. The only \ident{bind} case that generates the $\bnd{mem}(m, s)$
  node in the absence of let binding is~(2) and so $e=e_0.m(e_1, \ldots, e_n)$. From 
  (\emph{mem-val}), $p_0.m(p_1, \ldots, p_n) \rightsquigarrow p$ and from the compositionality 
  of $\rightsquigarrow$, it is the case that $e \rightsquigarrow p$.

\end{proof}

\begin{theorem}[Correctness]
(Same as let-free version, but using let elimination to say that this works for any expression)
\end{theorem}

\subsection{Reuse of previews}

* Inlining let leads to an equivalent dependency graph

* Given two expressions, if we inline lets and they are structurally different, then 
  their dependency graphs are structurally different
